An agent-based model for city networks based on interactions between firms

### 2020-03-04 main.tex
Update on Overleaf.

\begin{abstract}
    This paper introduces a generative model for urban networks defined by interactions between firms. The aim is to test several geographical properties emerging in such a network as metropolisation, internationalisation or regionalisation. The growth of the network is simulated by adding randomly new links defined by ownership between firms, with probabilities of occurring depending on the economic size of urban areas at origin and destination, industrial sector proximity between firms, the strength of links from the past as well as the geographical and socio-cultural distance. The simulation on a synthetic system of cities unveil phase transitions when changing interaction distance, and non-trivial patterns in the model behavior. Calibration of the model on real data on company ownership linkages for Europe in 2018 is also undertaken. Future work will consist in observing thanks to this model the impact of economic crisis on such networks.

@@

The use of network models to understand processes underlying the emergence and dynamics of such urban networks has already been widely tackled in the literature. \cite{dai2016simulating} propose a generative urban network model at the regional scale, combining geographical factors with network topological properties. Air transportation networks can also be understood as urban networks, and \cite{zanin2013modelling} reviews several models for such networks. Trade network have been extensively studied from the viewpoint of complex networks, but are generally driven at the country-level \cite{garlaschelli2005structure} because of scarce data at city-level. Input-output models, mostly used in Regional science to represent flows between geographical areas, are considered in some cases as a type of urban generative model \cite{jin1993generation}. In this stream of research, spatial interaction models \cite{dennett2013multilevel} can form the basis of urban network models for abstract flows \cite{dai2016generative} or physical infrastructure \cite{raimbault2020indirect}.

@@

The model is expected to capture a single macroscopic urban scale, i.e. the level of the urban system where basic entities are cities. Links are induced by underlying firms but these are not explicit. Cities are defined by an economic size, but also by an industrial sector profile. As \citep{martinus2018global} point out, a combination of several distances may play a role in establishing linkages. Geographical distance and industrial proximity and complementarity will typically be in concurrence \cite{cottineau2020nested}. Therefore, we combine several processes driving network growth: (i) geographical proximity (in our case the crow-fly distance, but it can be generalized to any type of accessibility); (ii) geopolitical proximity, which captures the higher propensity to establish links within existing submarkets despite the European common market (confirmed relevant by the fixed effect regression described above); (iii) city economic size, which has furthermore scaling laws properties; (iv) industrial proximity in terms of firm sector composition; (v) previous linkages (influence of history). This last point is crucial to capture accumulated effects and include path-dependency, which fits well the use of a simulation model instead of a more simple statistical model.

>>>>

\begin{abstract}
    This paper introduces a generative model for urban networks defined by interactions between firms. The aim is to test several geographical properties emerging in such a network as metropolisation, internationalisation and regionalisation and test the effect of specialisation on these phenomena. The growth of the network is simulated by adding randomly new links defined by ownership between firms, with probabilities of occurring depending on the economic size of urban areas at origin and destination, industrial sector proximity between firms, the strength of links from the past as well as the geographical and socio-cultural distance. The simulation on a synthetic system of cities unveil phase transitions when changing interaction distance, and non-trivial patterns in the model behavior. Calibration of the model on real data on company ownership linkages for Europe in 2018 is also undertaken. Future work will consist in observing thanks to this model the impact of economic crisis on such networks.

@@

The use of network models to understand processes underlying the emergence and dynamics of such urban networks has already been widely tackled in the literature. \cite{dai2016simulating} propose a generative urban network model at the regional scale, combining geographical factors with network topological properties. Air transportation networks can also be understood as urban networks, and \cite{zanin2013modelling} reviews several models for such networks. Trade network have been extensively studied from complex networks perspective, but are generally driven at the country-level \cite{garlaschelli2005structure} because of lack of data at city-level. Input-output models, mostly used in Regional science to represent flows between geographical areas, are considered in some cases as a type of urban generative model \cite{jin1993generation}. In this stream of research, spatial interaction models \cite{dennett2013multilevel} can form the basis of urban network models for abstract flows \cite{dai2016generative} or physical infrastructure \cite{raimbault2020indirect}.

@@

The model is expected to capture a single macroscopic urban scale, i.e. the level of the urban system where basic entities are cities. Links are induced by underlying firms but these are not explicit. Cities are defined by their economic size, but also by an industrial sector profile. As \citep{martinus2018global} point out, a combination of several distances may play a role in establishing linkages. Geographical distance, industrial proximity and complementarity will typically be in concurrence \cite{cottineau2020nested}. Therefore, we combine several processes driving network growth: (i) geographical proximity (in our case the crow-fly distance, but it can be generalized to any type of accessibility); (ii) geopolitical proximity, which captures the higher propensity to establish links within existing submarkets despite the European common market (confirmed relevant by the fixed effect regression described above); (iii) city economic size, which has furthermore scaling laws properties; (iv) industrial proximity in terms of firm sector composition; (v) previous linkages (influence of history). This last point is crucial to capture accumulated effects and include path-dependency, which fits well the use of a simulation model instead of a more simple statistical model.
